---
title: "Assignment 2"
output:
  word_document: default
  html_document:
    df_print: paged
---
Since credit datasets are non-numerical, we pick __applications__ datasets for regression analysis
export applications dataframes from Assignment 1

```{r}
library(ggplot2)
library(dplyr)
library(pastecs)
library(gridExtra)
library(psych)
library(data.table)
library(PerformanceAnalytics)
#data <- read.csv("applications_df.tsv")
```

### Summary of dataset
```{r}
str(data)
```

```{r}
`[`   <- function(...) {UseMethod("["  )}
`[<-` <- function(...) {UseMethod("[<-")}
`$`   <- function(...) {UseMethod("$"  )}
`$<-` <- function(...) {UseMethod("$<-")}

`[.function` <- `[.closure` <- function(func, arg_name) {
  if (arg_name %in% names(formals(func))) {
    formals(func)[[arg_name]]
  } else {
    warning("No such formal argument: ", arg_name)
  }
}

`[<-.function` <- `[<-.closure` <- function(func, arg_name, value) {
  fargs             <- formals(func)
  fargs[[arg_name]] <- value
  formals(func)     <- fargs
  func
}

`$.function` <- `$.closure` <- function(func, arg_name) {
  func[as.character(substitute(arg_name))]
  func
}

`$<-.function` <- `$<-.closure` <- function(func, arg_name, value) {
  func[as.character(substitute(arg_name))] <- value
  func
}
```


Clean data

```{r}
df <- data[c(3,7:13,19)]

df$AMT_INCOME_k = df$AMT_INCOME_TOTAL/1000
df$YRS_EMPLOYED <- round(abs(df$DAYS_EMPLOYED)/365, digits=2)
df$AGE <- round(abs(df$DAYS_BIRTH/365), digits=2)
df <- rename(df, FAMILY_SIZE = CNT_FAM_MEMBERS)
df$DAYS_EMPLOYED <- df$DAYS_BIRTH <- df$AMT_INCOME_TOTAL <- NULL
```

```{r}
summary(df)
```

### Generate models to predict an outcome

Linear Regression
```{r}
linear_mod <- lm(AMT_INCOME_k ~ AGE, data = df)
linear_mod
ggplot(linear_mod, aes(x = AGE, y = AMT_INCOME_k)) + geom_point() + ggtitle("Age vs Annual Income Regression model") + xlab("Age") + ylab("Income $k") 
#Unfortunately there is no linear relationship here

```

performance metrics
Coefficient of Determination
```{r}
coef(linear_mod)
exp(coef(linear_mod))
```
```{r}
linear_residplot <- function(linear_mod, nbreaks=10) {
  z <- rstudent(linear_mod)
  hist(z, breaks=nbreaks, freq=FALSE,
       xlab="Studentized Residual",
       main="Distribution of Errors")
  rug(jitter(z), col="brown")
  curve(dnorm(x, mean=mean(z), sd=sd(z)),
        add=TRUE, col="blue", lwd=2)
  lines(density(z)$x, density(z)$y,
        col="red", lwd=2, lty=2)
  legend("topright",
         legend = c( "Normal Curve", "Kernel Density Curve"),
         lty=1:2, col=c("blue","red"), cex=.7)
}

linear_residplot(linear_mod)
```



```{r}
library(lattice)
xyplot(AMT_INCOME_k ~ AGE, data = df,group = NAME_EDUCATION_TYPE, auto.key = list(title = "Annual Income vs Age by Education Levels", columns = 5))
```


Multiple Regression
```{r}
df.num = df[ , unlist(lapply(df, is.numeric))]

corr.test(df.num, use = "pairwise", method="pearson", adjust="none", alpha=.05)           # Can adjust p-values; see ?p.adjust for options
          
pairs(data=df,~ YRS_EMPLOYED +AGE + AMT_INCOME_k +FAMILY_SIZE)
chart.Correlation(df.num, method="pearson", histogram=TRUE, pch=20)
```


```{r}
## Model selection using the step function
model.null = lm(AMT_INCOME_k ~ 1, data= df)
model.null

```


```{r}
model.full = lm(AMT_INCOME_k ~ factor(CODE_GENDER) +factor(NAME_EDUCATION_TYPE) +YRS_EMPLOYED +AGE +FAMILY_SIZE, data = df)
summary(model.full)
```


```{r}
## Stepwise regressions
step(model.null,
     scope = list(upper=model.full),
     direction="both",
     data= df) 

```
```{r}
multi_mod = lm(AMT_INCOME_k ~ factor(NAME_EDUCATION_TYPE) +factor(CODE_GENDER) + YRS_EMPLOYED + AGE + FAMILY_SIZE, data = df)
summary(multi_mod)
```
```{r}
df$predy = predict(multi_mod)
plot(predy ~ AMT_INCOME_k,
     data= df,
     pch = 16,
     col = "darkgray",
     xlab="Actual total annual income",
     ylab="Predicted total annual income")
abline(0,1, col="blue", lwd=2)
```
```{r}
hist(residuals(multi_mod),
     col="lightblue")
```



### Explain the results of the models using performance metrics

Coefficient of Determination
```{r}
coef(multi_mod)
exp(coef(multi_mod))
```


plotting studentized residual
```{r}
residplot <- function(multi_mod, nbreaks=10) {
  z <- rstudent(multi_mod)
  hist(z, breaks=nbreaks, freq=FALSE,
       xlab="Studentized Residual",
       main="Distribution of Errors")
  rug(jitter(z), col="brown")
  curve(dnorm(x, mean=mean(z), sd=sd(z)),
        add=TRUE, col="blue", lwd=2)
  lines(density(z)$x, density(z)$y,
        col="red", lwd=2, lty=2)
  legend("topright",
         legend = c( "Normal Curve", "Kernel Density Curve"),
         lty=1:2, col=c("blue","red"), cex=.7)
}

residplot(multi_mod)

```

QQ plots- Assessing normality
```{r}
qqPlot(multi_mod, labels=row.names(df), id.method="identify",
       simulate=TRUE, main="Q-Q Plot")
```

### Improve the model
Assessing linearity
```{r}
library(car)
crPlots(multi_mod)
```

Assessing homoscedasticity
```{r}
ncvTest(multi_mod)
spreadLevelPlot(multi_mod)
```

Assessing multicollinearity
```{r}
vif(multi_mod) 
sqrt(vif(multi_mod)) > 2
```


